---
title: "Target sequencing DNA methylation report"
author: "Rafael Cuadrat"
output:
  html_notebook:
    toc:        TRUE
    toc_float:  TRUE
    theme:      "lumen"
    number_sections: FALSE
    code_folding: "hide"
    self_contained: TRUE
date: "`r Sys.Date()`"
---

# GOAL
Design probes for targeted sequencing of validation cohort by identifying the DMR overlap with disease associated regions and probes from the methylation atlas.

# Load Libraries
```{r Load_libraries}
library(methylKit)
library(tidyverse)
```

```{r Define_samples, results='asis'}
sample_ids_v = c(
 '2017',
 '2018', 
 '2019', 
 '2026', 
 '2027', 
 '2033',
 '3052',
 '3131',
 '3158', 
 'SK',
 'VF')
groups_v= c(1,1,2,3,3,2,1,2,1,0,0)
```

# Read in data
```{r Read_in_data}
p06_methyl_calls_v="/local/rcuadrat/cfdna_validation/out/06_methyl_calls/"
#samples<-read.csv("sample.txt",sep="\t")
```

Reading the methylcall logs and parsing bs ratios from the logs
```{r Parse_for_bs_conversion}
bs_ratio<-lapply(sample_ids_v, function(s) {
  FileInput = readLines(paste0(p06_methyl_calls_v,s,"_1_val_1_bt2.sorted.deduped_cpg_meth_calls.log")) 
  prot_pattern="average conversion rate = ";
  prot_string<-grep(prot_pattern,FileInput)
  BS_conversion<-FileInput[prot_string]
  BS_conversion<-str_split(BS_conversion," = ")[[1]][[2]]
  sample_id<-s
  data.frame(sample_id,BS_conversion)
  
})
bs_ratio_all<-do.call(rbind, bs_ratio)
```

Bisulfite converstion per sample (%):
```{r Plot_BS_conversion}
p<-ggplot(data=bs_ratio_all, aes(x=sample_id, y=as.numeric(BS_conversion))) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(y="Average conversion rate") + ylim(0, 100)
p
```

Read methylation files for each sample and get 
objects that only bases with coverage from all samples are retained.
```{r Read_methylation_files}
#functions
my.read.Tabix = function( dbpath,
                          dbtype,
                          sample.id, 
                          assembly ,
                          context,
                          resolution){
require(methylKit)
object.dt =data.table::fread( paste0('zcat ',dbpath))
methylKit:::.setMethylDBNames(object.dt)

num.records=Rsamtools::countTabix( dbpath )[[1]]

object <- new("methylRawDB",
          dbpath= dbpath,
          num.records=num.records,
          sample.id = sample.id, 
          assembly = assembly,
          context=context,
          resolution=resolution,
          dbtype=dbtype)
list(obj=object, 
     obj.dt=object.dt)

}
```

```{r Read_tabix_files}
# reading tabix files
 tabix.files.filtered.dt=lapply(sample_ids_v, function(s){
           dbpath = paste0(p06_methyl_calls_v, "methylKit/tabix_cpg/",s,"_1_val_1_bt2.sorted.deduped_cpg.txt.bgz" )
           tmp=my.read.Tabix(dbpath,
               "tabix",
                s, 
                "hg38" ,
                "CpG",
                "base")})
 names(tabix.files.filtered.dt) = sample_ids_v

 methRawList_fil=new("methylRawList",
   lapply(tabix.files.filtered.dt, function(x) x$obj),
   treatment=groups_v)
```

```{r Define_plot_output}
PLOTS="/local/rcuadrat/cfdna_validation/plots/"
```

Ploting methyl stats and cov stats
```{r Plot_stats, echo = TRUE, out.width="30%"}
tabix.files.mb = lapply(tabix.files.filtered.dt, function(x) x$obj)
 for(i in 1:length(tabix.files.mb)){
   png(paste0(PLOTS, 
     "getMethylationStats_filteredmincov3_",
     tabix.files.mb[[i]]@sample.id,
     "_mqc.png"))

     getMethylationStats(
       tabix.files.mb[[i]],
       plot=TRUE,
       both.strands=FALSE)

   dev.off()
 }

 for(i in 1:length(tabix.files.mb)){
   png(paste0(PLOTS, 
     "getCoverageStats_filteredmincov3_",
     tabix.files.mb[[i]]@sample.id,
     "_mqc.png"))

     getCoverageStats(
       tabix.files.mb[[i]],
       plot=TRUE,
       both.strands=FALSE)

   dev.off()
 }
```

Unifying a methylkit db object
```{r Unite}
methDB<-methylKit::unite(methRawList_fil)
```

Make PCA from raw DMRs per sample
```{r PCA_raw_DMRs}
PCASamples(methDB)
```

Convert to GRanges object
```{r Convert_GR}
gr_DMR_narrow=makeGRangesFromDataFrame(narrow_DMRs %>% separate(DMR,c("seqnames","start","end"),"\\."))
```

```{r Narrow_DMRs}
methDB_DMR_norrow<-regionCounts(methDB,gr_DMR_narrow)
methDB_DMR_norrow
```

```{r PCA_DMRs_narrow}
PCASamples(methDB_DMR_norrow)
```

Get the position of targeted CpGs
```{r Count_targeted_CpGs}
regions<-read.csv("/local/rcuadrat/cfdna_target/disease_tissue_spec_cpgs_hg38.txt",sep="\t")
print(paste0("We target a total of ",nrow(regions)," CpGs"))
print(paste0("However, we covered a total of ",nrow(as(methDB,"data.frame")), " CpGs in all the samples (=>5 reads)"))
```

```{r GR_regions}
regions_gr<-makeGRangesFromDataFrame(regions,seqnames.field=c("chr.hg38"),
                         start.field="position.hg38",
                         end.field=c("position.hg38"),
                         strand.field="strand.hg38")
```

Cpgs covered by all samples
```{r methDB_GR}
#maybe add extra cols
methDB_gr<-makeGRangesFromDataFrame(as(methDB,"data.frame"),keep.extra.columns = TRUE)
```

Overlap of actually sequenced cpgs and the targeted
```{r Find_overlap_seq_targeted}
#findOverlaps(methDB_gr,regions_gr)

overlap_sequenced_vs_targeted<-findOverlaps(methDB_gr,regions_gr,ignore.strand=TRUE)
overlap_sequenced_vs_targeted <-data.frame(overlap_sequenced_vs_targeted)
overlap_sequenced_vs_targeted_final<-data.frame(methDB_gr[overlap_sequenced_vs_targeted$queryHits])
NOT_overlap_sequenced_vs_targeted_final<-data.frame(methDB_gr[-overlap_sequenced_vs_targeted$queryHits])
print(paste0("From the ",nrow(regions)," targeted CpGs, there were ",nrow(overlap_sequenced_vs_targeted)," covered in all the samples"))
```

Do our non-covered CpGs overlap with simple tandem repeats ?
```{r Non_covered_cpg_tandem_rpts}
print(paste0("From the ",nrow(as(methDB,"data.frame"))," CpGs covered for all samples, there were ",nrow(NOT_overlap_sequenced_vs_targeted_final)," off target"))
```

Check how far from the target are all those many CpGs non-target but sequenced (NOT_overlap_sequenced_vs_targeted_final)
```{r Distance_non_target_cpgs}
cov_not_target<-NOT_overlap_sequenced_vs_targeted_final %>% dplyr::select(contains("coverage"))
names(cov_not_target)<-sample_ids_v
```

```{r Get_target_coverage}
cov_target<-overlap_sequenced_vs_targeted_final  %>% dplyr::select(contains("coverage"))
names(cov_target)<-sample_ids_v
```

Distribution of CpG coverage (number of reads covering the CpG)
```{r Plot_coverage_distribution}
ggplot(data=cov_target %>% gather() ,aes(value)) + 
  geom_histogram() + scale_x_continuous(lim=c(0,1000)) + 
    facet_wrap(~key, scales = 'free_x')
```

And then for the CpGs "out of target", those that were not in the BED file sent for the company: 
```{r Plot_cpg_out_target_distribution}
ggplot(data=cov_not_target  %>% gather(),aes(value)) + 
  geom_histogram() + scale_x_continuous(lim=c(0,1000))  +
    facet_wrap(~key, scales = 'free_x')
```

# Assign coordinates to targeted sequencing and check results
```{r Read_in_probes_methatlas}
probes = readRDS("~/illumina_probes_hg38_GRanges.RDS")
refExtended = read.csv("/local/agosdsc/projects/AAkalin_cfdTissue/results/extendedReference.csv")
illumina_probes<-as.data.frame(probes)
```

Get the coordinates from the CpGs on the meth_atlas
```{r Get_coordinates}
illumina_probes_atlas<-illumina_probes  %>% filter(ID %in% refExtended$CpGs)
```

Overlap sequenced cpgs and meth_atlas
```{r Overlap_seq_methatlas}
# findOverlaps(methDB_gr,makeGRangesFromDataFrame(illumina_probes_atlas))

tmp<-findOverlaps(methDB_gr,makeGRangesFromDataFrame(illumina_probes_atlas),ignore.strand=TRUE)
tmp <-data.frame(tmp)
tmp_bed3<-data.frame(methDB_gr[tmp$queryHits])
tmp_bed4<-data.frame(makeGRangesFromDataFrame(illumina_probes_atlas)[tmp$subjectHits])
final_overlaps<-merge(tmp_bed3,tmp_bed4,by=0)

```

Overlap between the "targeted cpgs" and cpgs from meth_atlas
```{r Find_overlaps}
#only gets "+" because there is only "+" on targeted regions_gr (/disease_tissue_spec_cpgs_hg38.txt)
tmp<-findOverlaps(makeGRangesFromDataFrame(illumina_probes_atlas),regions_gr,ignore.strand=TRUE)
tmp <-data.frame(tmp)
tmp_bed3<-data.frame(makeGRangesFromDataFrame(illumina_probes_atlas)[tmp$queryHits])
tmp_bed4<-data.frame(regions_gr[tmp$subjectHits])
final_overlaps<-merge(tmp_bed3,tmp_bed4,by=0)
```

```{r}
overlaps <- lapply(tabix.files.filtered.dt, function(x) {
  covered_cpgs<-(nrow(x$obj.dt))
  cov_cpgs_50<-(nrow(x$obj.dt %>% filter(coverage > 50)))
  sample1<-makeGRangesFromDataFrame(x$obj.dt,keep.extra.columns = TRUE)
  tmp<-findOverlaps(sample1,regions_gr,ignore.strand=TRUE)
  tmp <-data.frame(tmp)
  tmp_bed3<-data.frame(sample1[tmp$queryHits])
  tmp_bed4<-data.frame(regions_gr[tmp$subjectHits])
  final_overlaps<-merge(tmp_bed3,tmp_bed4,by=0)
  covered_targets<-(nrow(final_overlaps)/nrow(regions))*100
  
  
  sample1<-makeGRangesFromDataFrame(x$obj.dt%>% filter(coverage > 50),keep.extra.columns = TRUE)
  tmp<-findOverlaps(sample1,regions_gr,ignore.strand=TRUE)
  tmp <-data.frame(tmp)
  tmp_bed3<-data.frame(sample1[tmp$queryHits])
  tmp_bed4<-data.frame(regions_gr[tmp$subjectHits])
  final_overlaps<-merge(tmp_bed3,tmp_bed4,by=0)
  cov_targets_50<-(nrow(final_overlaps)/nrow(regions))*100
  
  sample_id<-x$obj@sample.id
  #methstat<-getMethylationStats(x$obj)
  data.frame(covered_cpgs,covered_targets,sample_id)


})
```

```{r}
cov_stats<-do.call(rbind, overlaps)
```

```{r}
#cov_stats<- merge(cov_stats,samples %>% dplyr::select(sample_id,Sample),by="sample_id") 

```

Distribution % of Targeted CpG covered by the sequencing (at least 5 reads) for each sample
```{r Plot_%_Covered_CpG_distribution}
p<-ggplot(data=cov_stats, aes(x=sample_id, y=covered_targets)) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(y="Target CpGs cov (%)") + ylim(0, 100)
p
```

Calculate methylation %
```{r Perc_Meth_Calculation}
new<- lapply(tabix.files.filtered.dt, function(x) {
  tmp<-x$obj.dt
  tmp$meth <-  (tmp$numCs / tmp$coverage) *100
  sample<-x$obj@sample.id
  #s<-samples %>% dplyr::filter(sample_id ==sample)
  #s<-samples
  #tmp$expected<-s$expected.meth
  #tmp$sample_id<-s$Sample
  #tmp$error <-  tmp$meth - tmp$expected
  tmp$sample <- sample
  tmp
 })
```

```{r New_stats_function}
new_stats<- lapply(new, function(x) {
  mean_methylation<-mean(x$meth)
  median_methylation<- median(x$meth)
  sample_id<-unique(x$sample)
  sample<-unique(x$sample_id)
  #expected_methylation<-unique(x$expected)
  data.frame(sample_id,mean_methylation,median_methylation)
 })
```

General stats: percentage of targeted covered, mean and median % methylation on CpGs, expected sample % CpG methylation 
```{r Final_stats}
new_stats_all<-do.call(rbind, new_stats)
final_stats<-merge(cov_stats,new_stats_all,by="sample_id")
final_stats <- final_stats %>% dplyr::select(-covered_cpgs) 
final_stats
```

```{r Testing}
#testing things
annotated_DMRs <- readRDS("/local/rcuadrat/cfdna_wgbs/tmp_dfs/annotated_DMRs.RDS")

ACS_STEMI_discovery_gr<-makeGRangesFromDataFrame(annotated_DMRs$ACS_Stemi)
```

Tiling the target/validation the same way the discovery was done 
```{r validation_cohort_tiling}
tiles.path=paste0("/local/rcuadrat/cfdna_validation/tileMethylCounts/")
tiles=tileMethylCounts(methDB, 
  win.size=500,
  step.size=500,
  save.db=TRUE,
  suffix="tiles_win500bp_step500bp",
  dbdir=tiles.path)
```

```{r set_names_all_samples}
sample_ids = c(
 'N1', # control
 'N2',
 'N3',
 'N4',
 'N5',
 'N6',
 'H26',
 'H28',

 'AC1', #stemi
 'AC2',
 'AC3',
 'AC4',
 'AC5',
 'AC6',
 'AC14',
 'AC15',

 'AC7',# nstemi
 'AC8',
 'AC9',
 'AC10',
 'AC11', 
 'AC12',
 'AC13',

 'AP1', #Acs/iAP
 'AP2',
 'AP3',
 'AP4',
 'AP5',  
 'AP6',

 'CS1', #CAD+posStressEcho
 'CS2',
 'CS3',
 'CS4', 
 'CS5',
 'CS6'
)



treatment=c(
    rep(0,8),  # control
    rep(1,8), #stemi
    rep(2,7),  # nstemi
    rep(3,6), #Acs/iAP
    rep(4,6) #CAD+posStressEcho
     )
#methT.path = '/local/AAkalin_cardiac/Results/cardiac/06_methyl_calls_bwameth/tileMethylCounts/methylBase_CpG_dT_tiles_win500bp_step500bp_123batch_v1.txt.bgz'
#meth.deT=methylKit:::readMethylBaseDB(methT.path,"tabix",
#                 sample_ids, "hg38" ,"CpG",
#                 "region",treatment,TRUE,skip=0)
```

```{r read_in_discovery_data}
p06_methyl_calls="/local/AAkalin_cardiac/Results/cardiac/06_methyl_calls_bwameth/"
tabix.files.filtered.dt_discovery=lapply(sample_ids, function(s){
           dbpath = paste0(p06_methyl_calls, s, "/tabix_CpG/",s,"_CpG_filtered.txt.bgz" )
           tmp=my.read.Tabix(dbpath,
               "tabix",
                s, 
                "hg38" ,
                "CpG",
                "base")})
names(tabix.files.filtered.dt_discovery) = sample_ids
```


```{r Merge_discovery_validation}
all<-c(tabix.files.filtered.dt,tabix.files.filtered.dt_discovery)
```

```{r Assign_groups}
methRawList_fil_all=new("methylRawList",
   lapply(all, function(x) x$obj),
   treatment=c(groups_v,treatment))
```

```{r Reduce_to_common_DMRs}
methDB_all<-methylKit::unite(methRawList_fil_all)
```

```{r Make_PCA}
PCASamples(methDB_all)
```
