---
title: "Ranking DMR based disgenet and stats"
author: "Rafael Cuadrat"
output:
  html_notebook:
    toc:        TRUE
    toc_float:  TRUE
    theme:      "lumen"
    number_sections: FALSE
    code_folding: "hide"
    self_contained: TRUE
date: "`r Sys.Date()`"
---

This notebook annotates DMRs for the 5 groups, adding gene info, GO, phenotype associations, DO and disgenet associations.  

Loading libraries and DMRs


```{r}
#.libPaths("~/newR")
```


```{r}
require(rGREAT)
library("annotatr")
library(GenomicRanges)
library(tidyverse)
library("biomaRt")
library(DOSE)
library(disgenet2r)
library(methylKit)
library(genomation)
library(pheatmap)
library(knitr)
library(kableExtra)
library(stringi)
library("enrichplot")
#library(dplyr)
library(gridExtra)



#OPEN DMR FILES

myDiff25p_tiled_list_v1 <- readRDS("/local/AAkalin_cardiac/Results/cardiac/RDS/myDiff25p_tiled_list_v1.RDS")

myDiff25p_tiled_list_v1_CAD_neg <-readRDS("/local/AAkalin_cardiac/Results/cardiac/RDS/myDiff25p_tiled_CAD+posStressEcho+negIntervention.vs.healthy.RDS")
myDiff25p_tiled_list_v1_CAD_pos <-readRDS("/local/AAkalin_cardiac/Results/cardiac/RDS/myDiff25p_tiled_CAD+posStressEcho+posIntervention.vs.healthy.RDS")

myDiff25p_tiled_list_v1[["CAD+posStressEcho+pos"]] <- myDiff25p_tiled_list_v1_CAD_pos
myDiff25p_tiled_list_v1[["CAD+posStressEcho+neg"]] <- myDiff25p_tiled_list_v1_CAD_neg

#coercing DMRs to a dataframes
DMR <- lapply(myDiff25p_tiled_list_v1, function(i){
  return(data.frame(i))
})
```


Annotate with rGREAT

```{r, echo = FALSE}
geneList <- lapply(myDiff25p_tiled_list_v1, function(dm) {
                                             job = submitGreatJob(
                                             dm,
                                             species = "hg38",
                                             includeCuratedRegDoms = TRUE,
                                             #bgChoice              = ifelse(is.null(bg), "wholeGenome", "data"),
                                             rule                  = c("basalPlusExt", "twoClosest", "oneClosest"),
                                             adv_upstream          = 5.0,
                                             adv_downstream        = 1.0,
                                             adv_span              = 1000.0,
                                             adv_twoDistance       = 1000.0,
                                             adv_oneDistance       = 1000.0,
                                             request_interval = 30,
                                             max_tries = 10,
                                             version = "default")
  print(availableOntologies(job))

  res = plotRegionGeneAssociationGraphs(job)
  return(list(job = job, res = res))
})

#coercing results do dataframe
geneList_df <- lapply(geneList, function(i){
  return(data.frame(i$res))
})
```


```{r}
liftOver.hg38_to_hg19 = function(mycur){
  require(rtracklayer)
  #require(liftOver) #Failed with error:  ‘‘liftOver’ is not a valid installed package’; #TODO: I got this error, that's why I use import.chain function below
  path = "/local/AAkalin_cardiac/Base/liftOver/hg38ToHg19.over.chain"
  ch = import.chain(path)
  seqlevelsStyle(mycur) = "UCSC"  # without it it doesn't work
  cur19 = liftOver(mycur, ch)
  cur19 = unlist(cur19)
  genome(cur19) = "hg19"
  return(cur19)
}
```


## reproducing Kasia results with old version of GREAT that included dropped ontologies (for example MSigDB)

```{r}
myDiff25p_tiled.hg19 = lapply(myDiff25p_tiled_list_v1, function(dm){
  liftOver.hg38_to_hg19(as(dm, "GRanges"))
})
```


```{r}
 great.output.list = lapply(myDiff25p_tiled.hg19, function(dm){
 job=submitGreatJob(dm, bg = NULL,
                      species               = "hg19",
                      includeCuratedRegDoms = TRUE,
                      #bgChoice              = ifelse(is.null(bg), "wholeGenome", "data"),
                      rule                  = c("basalPlusExt", "twoClosest", "oneClosest"),
                      adv_upstream          = 5.0,
                      adv_downstream        = 1.0,
                      adv_span              = 1000.0,
                      adv_twoDistance       = 1000.0,
                      adv_oneDistance       = 1000.0,
                      request_interval = 30,
                      max_tries = 10,
                      version = "3.0.0")
  print(availableOntologies(job))
  
 tb = getEnrichmentTables(job, ontology = c("GO Molecular Function" ,
                                              "GO Biological Process" ,
                                              "GO Cellular Component", 
                                              "BioCyc Pathway",
                                              "MSigDB Pathway",
                                              "MSigDB Oncogenic Signatures",
                                              "MSigDB Cancer Neighborhood",
                                              "Human Phenotype",
                                             "Disease Ontology",
                                              "Ensembl Genes"))
 list(job=job,tb= tb)
 })
```

```{r}
saveRDS(great.output.list,"~/tmp_dfs/GREAT_v3.RDS")

```


```{r}
saveRDS(geneList_df,"~/tmp_dfs/geneList_df.RDS")
saveRDS(geneList,"~/tmp_dfs/geneList.RDS")
```

```{r}
geneList_df<-readRDS("~/tmp_dfs/geneList_df.RDS")

```


Annoate with annotatR
The hg38_cpgs shortcut annotates regions to CpG islands, CpG shores, CpG shelves, and inter-CGI. 
The hg38_basicgenes shortcut annotates regions to 1-5Kb, promoters, 5’UTRs, exons, introns, and 3’UTRs.



```{r}
code = c('lncrna' = 'AH49559')
build_ah_annots(genome = 'hg38', ah_codes = code, annotation_class = 'custom')
print(annotatr_cache$list_env())


annots = c('hg38_cpgs', 'hg38_basicgenes', 'hg38_genes_intergenic',
           'hg38_genes_intronexonboundaries','hg38_enhancers_fantom','hg38_custom_lncrna') 


anno_lncrna = annotatr::build_annotations(genome = 'hg38', annotations = annots)



dm_annotated2 <- lapply(myDiff25p_tiled_list_v1, function(dm) {
    a=annotate_regions(
    regions = dm,
    annotations = anno_lncrna,
    ignore.strand = TRUE,
    quiet = FALSE)
  return(a)
})

```
```{r}
dm_annotated2_good <- dm_annotated2
```


```{r}
saveRDS(dm_annotated2, "~/tmp_dfs/annotR.RDS")
```

```{r}
dm_annotated2 <- readRDS("~/tmp_dfs/annotR.RDS")
```


```{r}
annot_ord<-summarize_annotations(dm_annotated2$ACS_Stemi)$annot.type

```



```{r,fig.width=10}

dm_vs_kg_annotations_STEMI = plot_annotation(
    annotated_regions = dm_annotated2$ACS_Stemi,
    annotation_order  = annot_ord,
    plot_title = '# of DMRs STEMI',
    x_label = ' Annotations',
    y_label = 'Count')

dm_vs_kg_annotations_NSTEMI = plot_annotation(
    annotated_regions = dm_annotated2$ACS_Nstemi,
    annotation_order  = annot_ord,
    plot_title = '# of DMRs NSTEMI',
    x_label = ' Annotations',
    y_label = 'Count')

dm_vs_kg_annotations_UA = plot_annotation(
    annotated_regions = dm_annotated2$ACS_iAP,
    annotation_order  = annot_ord,
    plot_title = '# of DMRs UA',
    x_label = '',
    y_label = 'Count')

```


```{r,fig.width=12,fig.height=10}
cowplot::plot_grid(dm_vs_kg_annotations_STEMI,dm_vs_kg_annotations_NSTEMI,dm_vs_kg_annotations_UA,labels="AUTO")

```

```{r}
#annot_STEMI<-summarize_annotations(dm_annotated2$ACS_Stemi)
#annot_STEMI$group<-"STEMI"
#annot_NSTEMI<-summarize_annotations(dm_annotated2$ACS_Nstemi)
#annot_NSTEMI$group<-"NSTEMI"
#annot_UA<-summarize_annotations(dm_annotated2$ACS_iAP)
#annot_UA$group<-"UA"
#annot_ACS<-rbind(annot_STEMI,annot_NSTEMI,annot_UA)
```

```{r}
#annot_ACS
```



```{r}
#annot_ACS <- annot_ACS %>% filter(annot.type == "hg38_custom_lncrna" | annot.type == "hg38_enhancers_fantom")
```



```{r}
#annot_ACS$annot.type <- gsub("hg38_custom_lncrna","lcRNA",annot_ACS$annot.type)
#annot_ACS$annot.type <- gsub("hg38_enhancers_fantom","Enhancers",annot_ACS$annot.type)

```


```{r,fig.width=10}
#ggplot(annot_ACS, aes(x=annot.type, y=n))+
#  geom_bar(stat='identity')+
#  ylab("Number of DRMs") +
#  xlab("") +
#  facet_wrap(~group)
```

```{r}

library(rtracklayer)
library(AnnotationHub)
library(Rsamtools)

```



Enhancers annotatiom from roadmap

```{r}
files <- list.files(path="roadmap", pattern="*.bed", full.names=TRUE, recursive=FALSE)
```




```{r}

bedfile <- lapply(files,function(x) {
  
  bedfile<-import(x, format="bed")
  bedfile<-bedfile[grepl("Enh", bedfile$name)]
  bedfile$sample<-str_split(x,"/")[[1]][[2]]
  return(bedfile)
  
})

all_bed<-do.call("c",bedfile)

```




```{r}
final_enhancer <- lapply(myDiff25p_tiled_list_v1, function(x) {
  tmp_bed2 <- findOverlaps(x, all_bed)
  tmp_bed2 <-data.frame(tmp_bed2)
  tmp_bed3<-data.frame(x[tmp_bed2$queryHits])
  tmp_bed4<-data.frame(all_bed[tmp_bed2$subjectHits])
  tmp_bed4 <- tmp_bed4 %>% dplyr::select(name,sample)
  final_enhancer_annotated<-merge(tmp_bed3,tmp_bed4,by=0)
  final_enhancer_annotated <-final_enhancer_annotated %>% group_by(seqnames,start,end,strand) %>% summarise_all(~paste(unique(stri_remove_empty(na.omit(.))), collapse = ','))
  
}
)


```





```{r}
ensembl=useMart("ensembl")
ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")

attributes = listAttributes(ensembl)



gene_ids_mart <- lapply(geneList_df, function(i){
        r <-getBM(attributes=c('hgnc_symbol','entrezgene_id'),
        filters = 'hgnc_symbol',
        values = i$gene,
        mart = ensembl,
        useCache = FALSE)
        return(r)

})

#adding entrez id on GREAT tables
geneList_entrez <- mapply(merge, x=geneList_df,y=gene_ids_mart, 
                          by.x="gene",by.y="hgnc_symbol",all.x = TRUE ,SIMPLIFY=FALSE) 


```

```{r}
saveRDS(geneList_entrez,"~/tmp_dfs/geneList_entrez.RDS")

```


```{r}
geneList_entrez<-readRDS("~/tmp_dfs/geneList_entrez.RDS")

```


Disgenet enrichment and plots
```{r}
enrichment_all <- lapply(geneList_entrez, function(i){
  enr<-enrichDGN(
    i$entrezgene_id,
    pvalueCutoff = 0.05,
    pAdjustMethod = "BH",
    minGSSize = 10,
    maxGSSize = 500,
    qvalueCutoff = 0.2,
    readable = FALSE
  )
  #enr@result <-enr@result %>% filter(str_detect(Description, regex(terms_for_regex, ignore_case = T)))
  plt<-dotplot(enr, showCategory=30)
  return(enr)
  
})
  
```

```{r}
dotplot(enrichment_all$ACS_Stemi,showCategory=30)
```


```{r}
saveRDS(enrichment_all,"~/tmp_dfs/enrichment_all.RDS")

```


```{r}
#enrichment_all<-readRDS("~/tmp_dfs/enrichment_all.RDS")

```


```{r}
# Get DMC specific to the disease   
#removing CAD group with both pos and neg myDiff25p_tiled_list_v1[-4]

myDiff25p_specdisease = lapply(1:length(myDiff25p_tiled_list_v1[-4]), 
							   function(i){
			myDiff25p_allothers=unlist(GRangesList(myDiff25p_tiled_list_v1[-4][-i]))
			fi=findOverlaps(myDiff25p_tiled_list_v1[-4][[i]], myDiff25p_allothers)
			diffmeth.l_list.i.unique = myDiff25p_tiled_list_v1[-4][[i]][-queryHits(fi),]
			diffmeth.l_list.i.unique
})

names(myDiff25p_specdisease)<-names(myDiff25p_tiled_list_v1[-4])

```


```{r}
# Get DMC specific to the disease   ONLY considering ACS

myDiff25p_specdisease_ACS = lapply(1:length(myDiff25p_tiled_list_v1[-4:-6]), 
							   function(i){
			myDiff25p_allothers=unlist(GRangesList(myDiff25p_tiled_list_v1[-4:-6][-i]))
			fi=findOverlaps(myDiff25p_tiled_list_v1[-4:-6][[i]], myDiff25p_allothers)
			diffmeth.l_list.i.unique = myDiff25p_tiled_list_v1[-4:-6][[i]][-queryHits(fi),]
			diffmeth.l_list.i.unique
})

names(myDiff25p_specdisease_ACS)<-names(myDiff25p_tiled_list_v1[-4:-6])
```



Merging DMR stats and GREAT annotations


```{r}
#Merging DMR stats and GREAT annotations
DMR_great <- mapply(left_join, DMR, geneList_df,
                    by.x=c("seqnames","start","end", "width", "strand"),by.y=c("seqnames","start","end", "width", "strand") ,SIMPLIFY=FALSE)


```



Get EFO associated genes from https://alpha.targetvalidation.org/

```{r}
EFO_CHD<-read.table("/home/rcastro/EFO_0001645-associated-diseases.tsv",sep="\t",header = TRUE)
EFO_CHD <-EFO_CHD %>% rename("gene"="symbol")
EFO_CHD <-EFO_CHD %>% dplyr::select(gene,overallAssociationScore) %>% rename("EFO0001645_score"="overallAssociationScore")


DMR_great_EFO <- lapply(DMR_great, function(dm) {

  
  df<-merge(dm,EFO_CHD,
            by="gene", all.x=TRUE ,SIMPLIFY=FALSE)
  return(df)
  
  })
  
```




Annotating with disgenet
Taking gene symbols 
Taking the diseases associated to each gene


```{r}
disgenet <- lapply(DMR_great, function(i){
  ge <- unique(i$gene)
  ge2<-split(ge, ceiling(seq_along(ge)/500))
  ge_chunks_ge2di<-lapply(ge2, function(ck) {
    ge2di <-gene2disease( gene = ck, verbose = TRUE, database="ALL")
    return(ge2di@qresult)
  })
  return(bind_rows(ge_chunks_ge2di))
  
})

#coercing results do dataframe
disgenet_df <- lapply(disgenet, function(i){
  return(data.frame(i))
})

#merging DMR table + disginet
DMR_great_disgenet <- mapply(merge, DMR_great_EFO, disgenet_df,
                      by.x=c("gene"),by.y=c("gene_symbol"), all.x=TRUE, SIMPLIFY=FALSE)
```

```{r}
saveRDS(DMR_great_disgenet,"/tmp_dfs/DMR_great_disgenet.RDS")

```

```{r}
DMR_great_disgenet<-readRDS("~/tmp_dfs/DMR_great_disgenet.RDS")


```



Get EFO vocabulary from disginet
```{r}
mapping_disgenet_EFO<-read.table(gzfile("/home/rcastro/disease_mappings.tsv.gz"),sep="\t", header = TRUE, quote = "\"")
mapping_disgenet_EFO <-mapping_disgenet_EFO %>% dplyr::filter(vocabulary == "EFO")
mapping_disgenet_EFO <-mapping_disgenet_EFO %>% rename("diseaseid" = "diseaseId")
mapping_disgenet_EFO <-mapping_disgenet_EFO %>% dplyr::select(diseaseid,vocabulary,code) %>% mutate(EFO_id=paste(vocabulary,code,sep="_")) %>% dplyr::select(diseaseid,EFO_id)

mapping_disgenet_EFO_2 <- lapply(DMR_great_disgenet, function(dm) {
  df<-merge(dm,mapping_disgenet_EFO,
            by="diseaseid", all.x=TRUE ,SIMPLIFY=FALSE)
  return(df)
})

```

Creating a column Y/N if finding terms related to heart diseases on disgenet gene->diseases associations (looking in disease name and disease class)
Summarizing a column with disease names and class (with heart related terms) (comma sepated terms)
Summarizing all the other columns (comma separated) (to make one row for one DMR)
Creating column with max values for scores of gene->associations (MAX_*). Those columns are only considering the scores from gene-diseases associations from heart related diseases.

```{r}
patent_genes <-c("AF161365", "HNRPF", "ACBD5", "TFCP2", "DDX18", "AF289562", "CD248", "CD79B", "CD19", "SPIB",
                 "BLK", "CD3D", "LCK", "TMC8", "CCT2", "S100A12", "MMP9", "CLEC4E", "ALOX5AP", "S100A8", "NAMPT", "RPL28",
                 "SSRP1", "AQP9", "GLT1D1", "NCF4", "NCF2", "CASP5", "H3F3B", "IL18RAP", "TXN", "TNFAIP6", "PLAUR", "IL8RB",
                 "BCL2A1", "TNFRSF10C", "PTAFR", "KCNE3", "LAMP2", "TLR4", "TYROBP", "SLAMF7", "CX3CR1", "KLRC4","CD8A")
```


```{r}
library(readxl)
circulation_MI <- read_excel("circulation_MI.xlsx")
circulation_CHD <- read_excel("circulation_CHD.xlsx")


circulation_genes_MI <-c(circulation_MI$Gene)
circulation_genes_CHD <-c(circulation_CHD$`Gene‡`)

```

```{r}
library(readr)
cluster8 <- read_table2("~/cluster8.csv", 
                        locale = locale(decimal_mark = ","))


cluster9 <- read_table2("~/cluster9.csv", 
                        locale = locale(decimal_mark = ","))

cluster10 <- read_table2("~/cluster10.csv", 
                        locale = locale(decimal_mark = ","))
```

```{r}
cluster8 <- cluster8$gene
cluster9 <- cluster9$gene
cluster10 <- cluster10$gene

```




```{r}

terms_for_regex="heart|coronary|atherosclerosis|inflammation|cardio|vascular|artery|infarction|myocard|cardiac|artery calcification|arterial stiffness"


na.strings=c("", "NA")

# string dectection for disases and sumarizing, creating Y|N column

heart_v2 <- lapply(mapping_disgenet_EFO_2, function(i){
  return( i %>%
            mutate(heart_disease_related = ifelse(  str_detect(disease_name, regex(terms_for_regex, ignore_case = T)) | 
                                              str_detect(disease_class_name,regex(terms_for_regex, ignore_case = T)),
                                              "Y","N" )) %>%
            tidyr::replace_na(list(heart_disease_related = "N")) %>%
            
            
            mutate(patent_genes = ifelse( gene %in% patent_genes  ,
                                              "Y","N" )) %>%
            tidyr::replace_na(list(patent_genes = "N")) %>%
            
            
            
            mutate(circulation_genes_MI = ifelse( gene %in% circulation_genes_MI  ,
                                              "Y","N" )) %>%
            tidyr::replace_na(list(circulation_genes_MI = "N")) %>%
            
            
            mutate(circulation_genes_CHD = ifelse( gene %in% circulation_genes_CHD  ,
                                              "Y","N" )) %>%
            tidyr::replace_na(list(circulation_genes_CHD = "N")) %>%
            
            
            mutate(cluster9 = ifelse( gene %in% cluster9  ,
                                              "Y","N" )) %>%
            tidyr::replace_na(list(cluster9 = "N")) %>%
            
            
                        mutate(cluster10 = ifelse( gene %in% cluster10  ,
                                              "Y","N" )) %>%
            tidyr::replace_na(list(cluster10 = "N")) %>%
            
            
                        mutate(cluster8 = ifelse( gene %in% cluster8  ,
                                              "Y","N" )) %>%
            tidyr::replace_na(list(cluster8 = "N")) %>%
            unique() %>%
            group_by(seqnames,start,end, width, strand,qvalue,meth.diff) %>% 
            summarize_all(~paste(unique(stri_remove_empty(na.omit(.))), collapse = ',')) %>%
            mutate(heart_disease_related = ifelse(str_detect(heart_disease_related, regex("Y")),"Y",heart_disease_related)) %>%
            mutate(patent_genes = ifelse(str_detect(patent_genes, regex("Y")),"Y",patent_genes)) %>%
            mutate(circulation_genes_MI = ifelse(str_detect(circulation_genes_MI, regex("Y")),"Y",circulation_genes_MI)) %>%
            mutate(circulation_genes_CHD = ifelse(str_detect(circulation_genes_CHD, regex("Y")),"Y",circulation_genes_CHD)) %>%
            mutate(cluster8 = ifelse(str_detect(cluster8, regex("Y")),"Y",cluster8)) %>% 
            mutate(cluster9 = ifelse(str_detect(cluster9, regex("Y")),"Y",cluster9)) %>%
            mutate(cluster10 = ifelse(str_detect(cluster10, regex("Y")),"Y",cluster10))
          
          
            )})

scores <- lapply(mapping_disgenet_EFO_2, function(i){
  return( i %>%
            filter( str_detect(disease_name, regex(terms_for_regex, ignore_case = T)) | 
                    str_detect(disease_class_name,regex(terms_for_regex, ignore_case = T)) ) %>%
            group_by(seqnames,start,end, width, strand,qvalue,meth.diff) %>% 
            summarise_at(c("score","ei","gene_dsi"), max, na.rm = TRUE) %>%
            rename(c('max_score'='score','max_ei'='ei','max_dsi'='gene_dsi')) %>% 
            mutate_all(function(x) ifelse(is.infinite(x), 0, x)) 
  )})

only_heart_col <- lapply(mapping_disgenet_EFO_2, function(i){
  return( i %>%
            filter( str_detect(disease_name, regex(terms_for_regex, ignore_case = T)) | 
                      str_detect(disease_class_name,regex(terms_for_regex, ignore_case = T)) ) %>%
            group_by(seqnames,start,end, width, strand,qvalue,meth.diff) %>% 
            summarise_at(c("disease_name","disease_class_name","EFO_id"),~paste(unique(stri_remove_empty(na.omit(.))), collapse = ', '), max, na.rm = TRUE) %>%
            rename(c('heart_disease_name'='disease_name','heart_disease_class_name'='disease_class_name','EFO_id_heart_related'='EFO_id'))

  )})

final_DMR_df <- mapply(left_join, x=heart_v2,y=scores, 
                       by.x=c("seqnames","start","end", "width", "strand"),by.y=c("seqnames","start","end", "width", "strand") ,SIMPLIFY=FALSE) 

final_DMR_df <- mapply(left_join, x=final_DMR_df,y=only_heart_col, 
                       by.x=c("seqnames","start","end", "width", "strand"),by.y=c("seqnames","start","end", "width", "strand") ,SIMPLIFY=FALSE)
                      
```

Calculating within group variation (SD)


```{r}
sample_ids = c('N1','N2','N3','N4','N5','N6','H26','H28',
  'AC1','AC2','AC3','AC4','AC5','AC6','AC14','AC15',
  'AC7','AC8','AC9','AC10','AC11','AC12','AC13',
  'AP1','AP2','AP3','AP4','AP5','AP6',
  'CS1','CS2','CS3','CS4','CS5','CS6')



treatment1=c(
    rep(0,8),  # control
    rep(1,8), #stemi
    rep(2,7),  # nstemi
    rep(3,6), #Acs/iAP
    rep(4,3), #CAD+posStressEcho+negIntervention
    rep(5,3) #CAD+posStressEcho+posIntervention
     )


treatment_descr=c(
  rep("Control",8),  # control
  rep("ACS_Stemi",8), #stemi
  rep("ACS_Nstemi",7),  # nstemi
  rep("ACS_iAP",6), #Acs/iAP
  rep("CAD+posStressEcho+neg",3),
  rep("CAD+posStressEcho+pos",3)
)

```


```{r}
#open methylDB with all tiles methylation

methT.path_tiled = '/local/AAkalin_cardiac/Results/cardiac/06_methyl_calls_bwameth/tileMethylCounts/methylBase_CpG_dT_tiles_win500bp_step500bp_123batch_v1.txt.bgz'
meth.deT_tiled=methylKit:::readMethylBaseDB(methT.path_tiled,"tabix",
                                     sample_ids, "hg38" ,"CpG",
                                     "region",treatment1,TRUE,skip=0)


#extracting methylation for all CpGs for all the samples
perc <- percMethylation(meth.deT_tiled)
perc <- data.frame(perc)
#defining groups
control<-c('N1','N2','N3','N4','N5','N6','H26','H28')
stemi<-c('AC1','AC2','AC3','AC4','AC5','AC6','AC14','AC15')
nstemi <-c( 'AC7','AC8','AC9','AC10','AC11','AC12','AC13')
iAP <- c('AP1', 'AP2','AP3','AP4','AP5','AP6')
cad_posStreessEcho <- c( 'CS1','CS2','CS3','CS4','CS5','CS6')
cad_posStreessEcho_neg <- c('CS1','CS2','CS3')
cad_posStreessEcho_pos <- c('CS4','CS5','CS6')

list_of_groups<-list(control,stemi,nstemi,iAP,cad_posStreessEcho,cad_posStreessEcho_pos,cad_posStreessEcho_neg)
names(list_of_groups) <- c("Control","ACS_Stemi","ACS_Nstemi","ACS_iAP","CAD+posStressEcho","CAD+posStressEcho+pos","CAD+posStressEcho+neg")

```



```{r}
#geting coordinates from methylation sites (percMethylation returned methylation % without coordinates)
coord <- read.table(gzfile('/local/AAkalin_cardiac/Results/cardiac/06_methyl_calls_bwameth/tileMethylCounts/methylBase_CpG_dT_tiles_win500bp_step500bp_123batch_v1.txt.bgz'),
                   sep="\t",comment.char = "#") 
coord <-coord %>% dplyr::select(c("V1","V2","V3","V4")) 
names(coord)<-c("seqnames","start","end", "strand")
#perc_meth<-merge(coord,perc,by=0)
```



```{r}

#calculating SD for perc methylation for all samples and for control
SDs <- lapply(list_of_groups, function(i){
  return(transform(perc, SD=apply(perc %>% dplyr::select(i),1, sd, na.rm = TRUE)) %>% dplyr::select(SD))
})

SD_df <- lapply(SDs[-1], function(i){
  df<-merge(coord,i,by=0)
  return(df %>% dplyr::select(-Row.names))
  })
SD_control<-merge(coord,SDs[1],by=0) %>% dplyr::select(-Row.names)
SD_control <- SD_control %>% rename("SD_control"="SD")
```


```{r}
#save SDs

saveRDS(SD_control,"~/tmp_dfs/SDs_control.RDS")
saveRDS(SD_df,"~/tmp_dfs/SDs_per_group.RDS")
```


```{r}
SD_control<-readRDS("~/tmp_dfs/SDs_control.RDS")
SD_df<-readRDS("~/tmp_dfs/SDs_per_group.RDS")


```


```{r}
#merging SDs with DMRs annotated
final_DMR_df_SDs <- mapply(left_join, x=final_DMR_df,y=SD_df, 
                       by.x=c("seqnames","start","end", "strand"),by.y=c("seqnames","start","end", "strand") ,SIMPLIFY=FALSE) 

final_DMR_df_SDs <- lapply(final_DMR_df_SDs,  function(i){
  
  df <- merge(i,SD_control,by.x=c("seqnames","start","end", "strand"),by.y=c("seqnames","start","end", "strand"),all.x = TRUE ,SIMPLIFY=FALSE)
  return(df)
})

```



```{r}
#adding info if the DMRs are disease exclusive
exc_df <- lapply(myDiff25p_specdisease, function(i){
  df<-data.frame(i) %>% mutate(exclusive ="Y")
 
  return(df %>% dplyr::select("seqnames","start","end", "width", "strand","exclusive"))
})


### merging disease exclusive col and REMOVING CAD group with both pos and neg
final_DMR_df_SDs <- mapply(left_join, x=final_DMR_df_SDs[-4],y=exc_df, 
                       by.x=c("seqnames","start","end", "width", "strand"),by.y=c("seqnames","start","end", "width", "strand") ,SIMPLIFY=FALSE)


```






Ranking by arrange and selecting columns for the final dataframe
Order of arrange absolute meth.diff (effect size) ---> qvalue -> SD (within group variation) ---> heart_disease_related (Y|N) --> max_score (max score of heart related diseases for that DMR) --> max_ei ---> max_dsi

```{r}
final_DMR_df_SDs_sorted <- lapply(final_DMR_df_SDs, function(i){
  
  return(arrange(i,desc(abs(meth.diff)),qvalue,SD,desc(heart_disease_related),desc(max_score),desc(max_ei),desc(max_dsi)) %>%
           dplyr::select("seqnames","start","end","width","strand","gene","distTSS","exclusive","patent_genes","circulation_genes_MI","circulation_genes_CHD",
                         "heart_disease_related","heart_disease_name","EFO_id_heart_related","heart_disease_class_name","max_score","EFO0001645_score",
                         "max_ei","max_dsi","meth.diff","qvalue","SD","SD_control") %>% 
           rename("heart_disease_related_disgenet"="heart_disease_related",
                                             "heart_disease_name_disgenet"="heart_disease_name",
                                             "EFO_id_heart_related_disgenet"="EFO_id_heart_related",
                                             "heart_disease_class_name_disgenet"="heart_disease_class_name",
                                             "max_score_disgenet"="max_score","EFO0001645_score_open_target"="EFO0001645_score"))
})



```





```{r}
#generate this table for displaying in the report_DMR.Rmd
a <- sapply(final_DMR_df_SDs_sorted,  function(i){
  
  df <- data.frame(i) %>%  dplyr::filter(heart_disease_related_disgenet == "Y")
  return(nrow(df))
})

b <- sapply(final_DMR_df_SDs_sorted,  function(i){
  
  df <- data.frame(i) %>%  dplyr::filter(heart_disease_related_disgenet == "Y" & exclusive == "Y")
  return(nrow(df))
})

counting<-data.frame(sapply(myDiff25p_tiled_list_v1[-4], length),
                     a,
                     sapply(myDiff25p_specdisease, length),
                     b)
names(counting)<-c("Number of DMRs","Number of DMRs with gene-disease associations","Number of DMRs disease exclusive","Number of DMRs disease exclusive with gene-disease associations")

```



Generate the sets for upset plot for all DMRs
```{r}

#this is for create a col in the DMR annotatead with info on witch groups the DMR is present
g1 <- final_DMR_df_SDs_sorted$ACS_Stemi %>% dplyr::select("seqnames","start","end") %>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g2 <- final_DMR_df_SDs_sorted$ACS_Nstemi %>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g3 <- final_DMR_df_SDs_sorted$ACS_iAP %>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g4 <- final_DMR_df_SDs_sorted$`CAD+posStressEcho+pos` %>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g5 <- final_DMR_df_SDs_sorted$`CAD+posStressEcho+neg` %>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))


g1$present_in_groups<-"ACS_Stemi"
g2$present_in_groups<-"ACS_Nstemi"
g3$present_in_groups<-"ACS_iAP"
g4$present_in_groups<-"CAD+posStressEcho+pos"
g5$present_in_groups<-"CAD+posStressEcho+neg"


####this is for veen diagram and upsetplot (for all DMRs)
v<-list(c(g1$dmr),c(g2$dmr),c(g3$dmr),c(g4$dmr),c(g5$dmr))

name_gs=c("Stemi","Nstemi","iAP","CAD_Echo_pos","CAD_Echo_neg")

names(v)<-name_gs


all_g<-rbind(g1,g2,g3,g4,g5)

all_g<-data.frame(all_g) 
all_g<- all_g %>% dplyr::select(seqnames,start,end,present_in_groups)
all_g <- all_g %>% group_by(seqnames,start,end) %>% summarize_all(~paste(unique(stri_remove_empty(na.omit(.))), collapse = ','))

final_DMR_df_SDs_sorted <- lapply(final_DMR_df_SDs_sorted,  function(i){
  
  df <- merge(i,all_g,by.x=c("seqnames","start","end"),by.y=c("seqnames","start","end"),all.x = TRUE ,SIMPLIFY=FALSE)
  return(df)
})

final_DMR_df_SDs_sorted <- lapply(final_DMR_df_SDs_sorted, function(i){
  
  return(arrange(i,desc(abs(meth.diff)),qvalue,SD,desc(heart_disease_related_disgenet),desc(max_score_disgenet),desc(max_ei),desc(max_dsi)))
})


```




DMRs for each disease/group vs. control

```{r,fig2, fig.height = 7, fig.width = 7}


###

# vennplot <-VennDiagram::venn.diagram(v,category.names = name_gs,output = TRUE ,
#           imagetype="png" ,filename = NULL,
#           resolution = 300,
#           cex=1.1,
#           cat.pos = c(30, 10, 200, 180,30),
#           fill = c("blue", "red","pink","green","grey"))
# grid.newpage()
# grid.draw(vennplot)

```



generate the sets for upset plot for heart_disease related DMRs

```{r}

g1_e <- final_DMR_df_SDs_sorted$ACS_Stemi %>%  dplyr::filter(heart_disease_related_disgenet == "Y" ) %>% dplyr::select("seqnames","start","end") %>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g2_e <- final_DMR_df_SDs_sorted$ACS_Nstemi %>%  dplyr::filter(heart_disease_related_disgenet == "Y"  ) %>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g3_e <- final_DMR_df_SDs_sorted$ACS_iAP %>%  dplyr::filter(heart_disease_related_disgenet == "Y" )%>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g4_e <- final_DMR_df_SDs_sorted$`CAD+posStressEcho+pos` %>%  dplyr::filter(heart_disease_related_disgenet == "Y" )%>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))
g5_e <- final_DMR_df_SDs_sorted$`CAD+posStressEcho+neg` %>%  dplyr::filter(heart_disease_related_disgenet == "Y")%>% dplyr::select("seqnames","start","end")%>% mutate(dmr = paste(seqnames,start,end,sep="_"))

v_e<-list(c(g1_e$dmr),c(g2_e$dmr),c(g3_e$dmr),c(g4_e$dmr),c(g5_e$dmr))

name_gs_e=c("Stemi","Nstemi","iAP","CAD_Echo_pos","CAD_Echo_neg")

names(v_e)<-name_gs_e

UpSetR::upset(UpSetR::fromList(v_e), order.by = "freq")

```

```{r}
#Just for ACS groups with genes
v_ACS<-list(c(g1$dmr),c(g2$dmr),c(g3$dmr))

name_gs_e_ACS=c("Stemi","Nstemi","UA")

names(v_ACS)<-name_gs_e_ACS

UpSetR::upset(UpSetR::fromList(v_ACS), order.by = "freq")
```





```{r}

#Just for ACS groups with genes
v_e_ACS<-list(c(g1_e$dmr),c(g2_e$dmr),c(g3_e$dmr))

name_gs_e_ACS=c("Stemi","Nstemi","iAP")

names(v_e_ACS)<-name_gs_e_ACS

UpSetR::upset(UpSetR::fromList(v_e_ACS), order.by = "freq")

```





```{r}
saveRDS(final_DMR_df_SDs_sorted,"/local/AAkalin_cardiac/Results/cardiac/ranking/annotated_ranked_DMRs.RDS")
```


```{r}
final_DMR_df_SDs_sorted<-readRDS("/local/AAkalin_cardiac/Results/cardiac/ranking/annotated_ranked_DMRs.RDS")
```




Here we join info from the CpG atlas for tissue specific CpGs

```{r}
#matrix with altas reference
extendedReference <- read_csv("/local/agosdsc/projects/AAkalin_cfdTissue/results/extendedReference.csv")
#location of the cpgs 
cpg_tissuespec=data.table::fread("/local/agosdsc/projects/AAkalin_cfdTissue/results/extendedReference_probeRegionInfo.csv")
#get only hg38 coordinates
cpg_tissuespec=cpg_tissuespec[-which(cpg_tissuespec$CHR_hg38==""),]
#renaming the cols
cpg_tissuespec<-cpg_tissuespec %>% dplyr::select(CpG,CHR_hg38,Start_hg38,End_hg38,Strand_hg38) %>% rename("chr"="CHR_hg38","start"="Start_hg38","end"="End_hg38","strand"="Strand_hg38")

```


```{r}


```

```{r}

#getting the extreme values (biggest distance from the avg) and the tissue/cell that has this extreme values (cell/tissue specific cpgs)
extendedReference <-extendedReference %>% remove_rownames %>% column_to_rownames(var="CpGs")
extendedReference$avg <- rowMeans(extendedReference)
extendedReference$max <-colnames(extendedReference[,1:28])[apply(extendedReference[,1:28], 1, which.max)]
extendedReference$min <-colnames(extendedReference[,1:28])[apply(extendedReference[,1:28], 1, which.min)]
extendedReference$max_value<-apply(X=extendedReference[,1:28], MARGIN=1, FUN=max)
extendedReference$min_value<-apply(X=extendedReference[,1:28], MARGIN=1, FUN=min)
extendedReference$max_mins_avg<- extendedReference$max_value - extendedReference$avg
extendedReference$avg_mins_min<- extendedReference$avg - extendedReference$min_value
extendedReference$specific <- ifelse(extendedReference$max_mins_avg >extendedReference$min_value, extendedReference$max, extendedReference$min)
cpgs_tissues<-extendedReference %>% dplyr::select(specific)

#merging the dataframe indicating CpGs --> tissue
cpgs_tissues <-tibble::rownames_to_column(cpgs_tissues, "CpG")
cpg_tissuespec <-merge(cpg_tissuespec,cpgs_tissues,by="CpG")
cpg_tissuespec_gr=makeGRangesFromDataFrame(cpg_tissuespec,
              keep.extra.columns=TRUE,
              ignore.strand=FALSE,       
              seqnames.field="chr",   
              start.field="start",    
              end.field="end",    
              strand.field="strand",   
              starts.in.df.are.0based=TRUE) 

cpg_tissuespec_grF = cpg_tissuespec_gr[strand(cpg_tissuespec_gr)=="+"]
cpg_tissuespec_grR = cpg_tissuespec_gr[strand(cpg_tissuespec_gr)=="-"]

cpg_tissuespec_grR = GRanges(seqnames(cpg_tissuespec_grR),
							 IRanges(start(cpg_tissuespec_grR)-1L,
							 		 end(cpg_tissuespec_grR)-1L),
							 "+",mcols=data.frame(cpg_tissuespec_grR) %>% dplyr::select(CpG,specific))
names(values(cpg_tissuespec_grR)) <-c("CpG","specific")

cpg_tissuespec_gr = sort(c(cpg_tissuespec_grF,cpg_tissuespec_grR))
end(cpg_tissuespec_gr) = end(cpg_tissuespec_gr)-1


list_treat<-unique(names(myDiff25p_tiled_list_v1[-4]))

#making GRanges from the final_DMR_df_SDs_sorted to merge with CpG atlas tissue specific info
final_cpgs_specifics <- lapply(list_treat, function(i){
  tmp_annotated<-makeGRangesFromDataFrame(final_DMR_df_SDs_sorted[[i]],keep.extra.columns = TRUE)
    tmp_annotated2<-findOverlaps(tmp_annotated,
                                 cpg_tissuespec_gr)
    tmp_annotated2 <-data.frame(tmp_annotated2)
    tmp_annotated3<-data.frame(tmp_annotated[tmp_annotated2$queryHits])
    tmp_annotated4<-data.frame(cpg_tissuespec_gr[tmp_annotated2$subjectHits])
    tmp_annotated4 <- tmp_annotated4 %>% dplyr::select(CpG,specific)
    final_tmp<-merge(tmp_annotated3,tmp_annotated4,by=0)
    final_tmp <-final_tmp %>% group_by(seqnames,start,end,strand) %>% summarise_all(~paste(unique(stri_remove_empty(na.omit(.))), collapse = ','))
    final_tmp <- final_tmp %>% dplyr::select("seqnames","start" ,"end","strand","CpG","specific")
  
  return(final_tmp)
})
names(final_cpgs_specifics) <-list_treat

```


```{r}


#adding CpG tissue specific annotation to the sorted-annotated DMRs

final_DMR_df_SDs_sorted <- mapply(left_join, x=final_DMR_df_SDs_sorted,y=final_cpgs_specifics, 
                       by.x=c("seqnames","start","end", "width", "strand"),by.y=c("seqnames","start","end", "width", "strand") ,SIMPLIFY=FALSE)


```




Here we start to look to DMRs without perc filter, checking if some of them has CpGs tissue specific for interesting tissues/cell-types

```{r}
#looking for tissues of interest
tissue_terms="Monocytes_EPIC|Left_atrium|coronary_artery|right_atrium_auricular_region|heart_left_ventricle|Left_atrium|Vascular_endothelial_cells"
```



```{r}
#opening the methylation call for all tiles and getting percMethyl info

#methT.path_tiled = '/local/AAkalin_cardiac/Results/cardiac/06_methyl_calls_bwameth/tileMethylCounts/methylBase_CpG_dT_tiles_win500bp_step500bp_123batch_v1.txt.bgz'
#meth.deT_tiled=methylKit:::readMethylBaseDB(methT.path_tiled,"tabix",
#                                     sample_ids, "hg38" ,"CpG",
#                                     "region",treatment1,TRUE,skip=0)


meth.deT_tiled_gr<-as(meth.deT_tiled,"GRanges")




percMethylation_batch123_v1 <- readRDS("/local/AAkalin_cardiac/Results/cardiac/RDS/percMethylation_batch123_v1.RDS")

values(meth.deT_tiled_gr)<-percMethylation_batch123_v1
```



```{r}

#metadata table
TBL = data.frame(sample_ids=sample_ids,
                 treatment_descr=treatment_descr,
                 treatment=treatment1,
                 stringsAsFactors = FALSE)
```


```{r}
#calling diffmeth (probably I can just replace this for opening the files Kasia generated)
for(treat in unique(TBL$treatment_descr)[-1]){

  TBL.subset=TBL[TBL$treatment_descr %in% c("Control", treat),]
  print(TBL.subset)
  meth.deT_tiled.subset =reorganize(meth.deT_tiled,
                   sample.ids=TBL.subset$sample_ids,
                   treatment=TBL.subset$treatment,
                   save.db = FALSE )

  myDiff=calculateDiffMeth(meth.deT_tiled.subset,
  chunk.size = 1e+05,
  save.db = TRUE,
  mc.cores=40,
  suffix=paste0("tiles_win500bp_step500bp_batch123_v1",
                "_treatment",
                stringr::str_replace_all(treat, "[[:punct:]]", "_")),
  overdispersion="MN",
  test="Chisq",
  dbdir="~/test"
  )
}
```


```{r}
#opening the files saved in the diffmeth call
 dmc_treat =
 lapply(unique(TBL$treatment_descr)[-1],
   function(treat){

     TBL.subset=TBL[TBL$treatment_descr %in% c("Control", treat),]
     dbpath=paste0("~/test/",
                           paste0("methylDiff_tiles_win500bp_step500bp_batch123_v1",
                                   "_treatment",
                                   treat,
                                   #stringr::str_replace_all(treat, "[[:punct:]]", ""),
                                   ".txt.bgz"))
     methylKit:::readMethylDiffDB(dbpath,
                      "tabix",
                      TBL.subset$sample_ids,
                      "hg38",
                      "CpG",
                      "region",
                      TBL.subset$treatment,
                      TRUE,
                      skip=0)
   })

#giving names to the itens in the list
names(dmc_treat)<-unique(TBL$treatment_descr)[-1]


#overlap DMRs not filtered and cpgs tissue specific

methyl_perc_atlas_cpg_regions<-findOverlaps(cpg_tissuespec_gr, meth.deT_tiled_gr)

hits_all_meth<-meth.deT_tiled_gr[data.frame(methyl_perc_atlas_cpg_regions)$subjectHits]

hits_all_tissue<-cpg_tissuespec_gr[data.frame(methyl_perc_atlas_cpg_regions)$queryHits]

hits_all_tissue <- data.frame(hits_all_tissue) %>% dplyr::select(CpG,specific) 



hits_all_meth<-merge(hits_all_meth,hits_all_tissue,by=0) %>%  filter( str_detect(specific, regex(tissue_terms, ignore_case = T)))
hits_all_meth <- hits_all_meth %>% dplyr::select(-Row.names)

cols=names(hits_all_meth %>% dplyr::select(-CpG,-specific))

hits_all_meth <-hits_all_meth %>% group_by_at(cols) %>% summarise_all(~paste(unique(stringi::stri_remove_empty(na.omit(.))), collapse = ','))

dmc_treat_gr =
 lapply(dmc_treat,function(i)
{
  as(i,"GRanges")
})

dmc_treat_df =
 lapply(dmc_treat_gr,function(i)
{
  tmp<-data.frame(i)
})

#adding names of groups to stats cols to merge 
dmc_treat_cpg_altas<- lapply(names(dmc_treat_df),function(i)
{

  tmp<-rename_with(dmc_treat_df[[i]], .fn = ~paste0(i,"_","qvalue"), .cols = qvalue )
  tmp<-rename_with(tmp, .fn = ~paste0(i,"_","pvalue"), .cols = pvalue )
  tmp<-rename_with(tmp, .fn = ~paste0(i,"_","meth.diff"), .cols = meth.diff )
  data.frame(left_join(x=hits_all_meth,y=tmp,by.x=c("seqnames","start","end","width","strand"),by.y=c("seqnames","start","end","width","strand")))

})

names(dmc_treat_cpg_altas)<-names(dmc_treat_df)


#merging all the DMRs from all groups
tmp1<-merge(dmc_treat_cpg_altas$ACS_Stemi,dmc_treat_cpg_altas$ACS_Nstemi %>% dplyr::select(ACS_Nstemi_pvalue , ACS_Nstemi_qvalue , ACS_Nstemi_meth.diff ),by=0,sort = FALSE) %>%
  dplyr::select(-Row.names)
tmp1<-merge(tmp1,dmc_treat_cpg_altas$ACS_iAP %>% dplyr::select(ACS_iAP_pvalue , ACS_iAP_qvalue , ACS_iAP_meth.diff ),by=0,sort = FALSE) %>%
  dplyr::select(-Row.names)
tmp1<-merge(tmp1,dmc_treat_cpg_altas$`CAD+posStressEcho+neg` %>% 
              dplyr::select("CAD.posStressEcho.neg_pvalue" , "CAD.posStressEcho.neg_qvalue"  , "CAD.posStressEcho.neg_meth.diff" ),by=0,sort = FALSE) %>%
  dplyr::select(-Row.names)

tmp1<-merge(tmp1,dmc_treat_cpg_altas$`CAD+posStressEcho+pos` %>% 
              dplyr::select("CAD.posStressEcho.pos_pvalue" , "CAD.posStressEcho.pos_qvalue"  , "CAD.posStressEcho.pos_meth.diff" ),by=0,sort = FALSE) %>%
  dplyr::select(-Row.names)

```



Generate PDF for all regions with tissue of interest without any filter 
```{r}
pdf(paste("MR_with_altas_cpg",".pdf",sep=""),height=8)
for (i in seq(1,nrow(tmp1))) {
  
    ##creating a grid
    grid.newpage()
    # 2x2 layout
    pushViewport(viewport(layout=grid.layout(10, 1)))
    MRs <- tmp1[i,]
    names(MRs)<-str_replace(names(MRs),".posStressEcho.","++")
    #MRs_gr <-as(MRs,"GRanges")
    forplot<-data.frame(MRs) %>% dplyr::select(sample_ids)
    forplot<-data.frame(t(forplot),treatment_descr)
    names(forplot)<- c("meth","treatment_descr")
    forplot$treatment_descr <- str_replace(forplot$treatment_descr,"posStressEcho","")
    forplot$treatment_descr <- factor(forplot$treatment_descr,levels=unique(forplot$treatment_descr))
    
    vplayout <- viewport(layout.pos.col=1,layout.pos.row=1:7)
    
    chr <- MRs$seqnames
    #genome coordinates for the DMR
    start<-MRs$start
    end<-MRs$end
    title=paste(chr,start,end,sep=" ")
    #
    p1<-ggplot(forplot, aes(x=treatment_descr, y=as.numeric(meth))) + 
      geom_boxplot(alpha=0.9) + 
      geom_jitter(shape=16, position=position_jitter(0.2)) + 
      scale_fill_manual(values=c("grey","red")) +
      xlab("") +
      ylab("% methylation") +
      theme(legend.position="none") + 
      ggtitle(paste(MRs$specific,title=" ")) +
      theme(plot.title = element_text(size=13,hjust = 0.5)) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      panel.background = element_blank())
      
    print(p1,vp=vplayout)
    
    vplayout2 <- viewport(layout.pos.col=1,layout.pos.row=8)
    tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),base_size = 8)
    grid.table(MRs %>% dplyr::select(contains("qvalue")),vp=vplayout2,theme=tt)
    
    vplayout3 <- viewport(layout.pos.col=1,layout.pos.row=9)
    tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),base_size = 8)
    grid.table(MRs %>% dplyr::select(contains("diff")),vp=vplayout3,theme=tt)

    vplayout4 <- viewport(layout.pos.col=1,layout.pos.row=10)
    tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),base_size = 8)
    grid.table(MRs %>% dplyr::select(seqnames,   start ,    end, CpG, specific),vp=vplayout4,theme=tt)
    popViewport()

}
```


```{r}
saveRDS(tmp1,"MR_with_atlas_cpg.RDS")
```

Filtering by significance in at least one group

```{r}
tmp1_sig <- tmp1 %>% dplyr::filter(ACS_Nstemi_qvalue < 0.01 | ACS_Stemi_qvalue < 0.01 | ACS_iAP_qvalue  < 0.01 | CAD.posStressEcho.neg_qvalue < 0.01 | CAD.posStressEcho.pos_qvalue < 0.01 )
```

```{r}
saveRDS(tmp1_sig,"MR_with_atlas_cpg_sig.RDS")

```


```{r}
pdf(paste("MR_with_altas_cpg_sig",".pdf",sep=""),height=8)
for (i in seq(1,nrow(tmp1_sig))) {
  
    ##creating a grid
    grid.newpage()
    # 2x2 layout
    pushViewport(viewport(layout=grid.layout(10, 1)))
    MRs <- tmp1_sig[i,]
    names(MRs)<-str_replace(names(MRs),".posStressEcho.","++")
    #MRs_gr <-as(MRs,"GRanges")
    forplot<-data.frame(MRs) %>% dplyr::select(sample_ids)
    forplot<-data.frame(t(forplot),treatment_descr)
    names(forplot)<- c("meth","treatment_descr")
    forplot$treatment_descr <- str_replace(forplot$treatment_descr,"posStressEcho","")
    forplot$treatment_descr <- factor(forplot$treatment_descr,levels=unique(forplot$treatment_descr))
    
    vplayout <- viewport(layout.pos.col=1,layout.pos.row=1:7)
    
    chr <- MRs$seqnames
    #genome coordinates for the DMR
    start<-MRs$start
    end<-MRs$end
    title=paste(chr,start,end,sep=" ")
    #
    p1<-ggplot(forplot, aes(x=treatment_descr, y=as.numeric(meth))) + 
      geom_boxplot(alpha=0.9) + 
      geom_jitter(shape=16, position=position_jitter(0.2)) + 
      scale_fill_manual(values=c("grey","red")) +
      xlab("") +
      ylab("% methylation") +
      theme(legend.position="none") + 
      ggtitle(paste(MRs$specific,title,sep=" ")) +
      theme(plot.title = element_text(size=13,hjust = 0.5)) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      panel.background = element_blank())
      
    print(p1,vp=vplayout)
    
    vplayout2 <- viewport(layout.pos.col=1,layout.pos.row=8)
    tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),base_size = 8)
    grid.table(MRs %>% dplyr::select(contains("qvalue")),vp=vplayout2,theme=tt)
    
    vplayout3 <- viewport(layout.pos.col=1,layout.pos.row=9)
    tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),base_size = 8)
    grid.table(MRs %>% dplyr::select(contains("diff")),vp=vplayout3,theme=tt)

    vplayout4 <- viewport(layout.pos.col=1,layout.pos.row=10)
    tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),base_size = 8)
    grid.table(MRs %>% dplyr::select(seqnames,   start ,    end, CpG, specific),vp=vplayout4,theme=tt)
    popViewport()

}
```




Checking if DMRs have cpgs from Horvath list of biological clock cpgs

```{r}
library(cgageR)
data("HorvathLongCGlist")
```



```{r}
manifest<-read.csv(gzfile("/local/AAkalin_cardiac/Base/Annotation/hg38/HM450.hg38.manifest.tsv.gz"),sep = '\t')
manifest<-manifest[manifest$probeID %in% c(as.character( HorvathLongCGlist$MR_var)),]
manifest_GR<-makeGRangesFromDataFrame(manifest,keep.extra.columns = TRUE,seqnames.field="CpG_chrm",start.field="CpG_beg",end.field="CpG_end",
                         strand.field="probe_strand",starts.in.df.are.0based=TRUE)

```



```{r}

#making GRanges from the final_DMR_df_SDs_sorted to merge with CpG atlas tissue specific info
final_cpgs_age <- lapply(list_treat, function(i){
  tmp_annotated<-makeGRangesFromDataFrame(final_DMR_df_SDs_sorted[[i]],keep.extra.columns = TRUE)
    tmp_annotated2<-findOverlaps(tmp_annotated,
                                 manifest_GR)
    tmp_annotated2 <-data.frame(tmp_annotated2)
    tmp_annotated3<-data.frame(tmp_annotated[tmp_annotated2$queryHits])
    tmp_annotated4<-data.frame(cpg_tissuespec_gr[tmp_annotated2$subjectHits])
    tmp_annotated4 <- tmp_annotated4 %>% dplyr::select(CpG,specific)
    final_tmp<-merge(tmp_annotated3,tmp_annotated4,by=0)
    final_tmp <-final_tmp %>% group_by(seqnames,start,end,strand) %>% summarise_all(~paste(unique(stri_remove_empty(na.omit(.))), collapse = ','))
    #final_tmp <- final_tmp %>% dplyr::select("seqnames","start" ,"end","strand","CpG","specific")
  
  return(final_tmp)
})
names(final_cpgs_age) <-list_treat
```






```{r}
sessionInfo()
```
