---
title: "R Notebook"
output: html_notebook
---

# GOAL
The goal of this script is to predict the cell/tissue-type proportions via NNLS modelling where ccfDNA originates from.

# Load Libraries
```{r Load_libraries}
library(tidyverse)
library(deconvR)
```

# Read in Data
```{r Read_in_data}
probes = readRDS("~/illumina_probes_hg38_GRanges.RDS")
refExtended = read.csv("/local/agosdsc/projects/AAkalin_cfdTissue/results/extendedReference.csv")
suppl_tbl_2<-readRDS("/local/rcuadrat/cfdna_wgbs/suppl_tbl_2.RDS")

```

# Set theme
```{r Set_Theme_for_Plots}
theme_Publication <- function(base_size=14, base_family="sans") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5, margin = margin(0,0,20,0)),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line.x = element_line(colour="black"),
               axis.line.y = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.box = "vetical",
               legend.key.size= unit(0.5, "cm"),
               #legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
       ))
      
}
```

Make metadata table
```{r Make_metadata_table}
sample_ids = c(
 'N1', # control
 'N2',
 'N3',
 'N4',
 'N5',
 'N6',
 'H26',
 'H28',
 'AC1', #stemi
 'AC2',
 'AC3',
 'AC4',
 'AC5',
 'AC6',
 'AC14',
 'AC15',
 'AC7',# nstemi
 'AC8',
 'AC9',
 'AC10',
 'AC11', 
 'AC12',
 'AC13',
 'AP1', #Acs/iAP
 'AP2',
 'AP3',
 'AP4',
 'AP5',  
 'AP6',
 'CS1', # CAD+posStressEcho+negIntervention
 'CS2',
 'CS3',
 'CS4', # CAD+posStressEcho+posIntervention
 'CS5',
 'CS6'
)
treatment1=c(
    rep(0,8),  # control
    rep(1,8), #stemi
    rep(2,7),  # nstemi
    rep(3,6), #Acs/iAP
    rep(4,3), #CAD+posStressEcho+negIntervention
    rep(5,3) #CAD+posStressEcho+posIntervention
     )


treatment_descr=c(
  rep("Control",8),  # control
  rep("ACS_Stemi",8), #stemi
  rep("ACS_Nstemi",7),  # nstemi
  rep("ACS_iAP",6), #Acs/iAP
  rep("CAD+posStressEcho+neg",3),
  rep("CAD+posStressEcho+pos",3)
)

methT.path = '/local/AAkalin_cardiac/Results/cardiac/06_methyl_calls_bwameth/tileMethylCounts/methylBase_CpG_dT_tiles_win500bp_step500bp_123batch_v1.txt.bgz'
meth.deT=methylKit:::readMethylBaseDB(methT.path,"tabix",
                                      sample_ids, "hg38" ,"CpG",
                                      "region",treatment =  treatment1,TRUE,skip=0)

desc<-data.frame(treatment1,treatment_descr,sample_ids)
desc<-desc %>% rename("treatment"="treatment1")
desc<-unique(desc)
desc$treatment_descr <- gsub("iAP","UA",desc$treatment_descr)
```

Mapping whole-genome bisulfite sequencing data to their probe IDs
```{r Map_data_to_ids}
meth.deT@treatment=treatment1
mapping = deconvR::BSmeth2Probe(probe_id_locations = probes, WGBS_data = meth.deT)

```

# Cell type proportion prediction

Prediction of cell-type proportions of bulk methylation profiles, as well as partial R-squared values for each sample using non-negative least square regression.
```{r Cell_type_prediction}
decon_nnls_extended = deconvR::deconvolute(reference=refExtended, bulk=mapping, model="nnls")

decon_nnls_extended_0 = decon_nnls_extended[1:8,]    
decon_nnls_extended_1 = decon_nnls_extended[9:16,]
decon_nnls_extended_2 = decon_nnls_extended[17:23,]
decon_nnls_extended_3 = decon_nnls_extended[24:29,]
decon_nnls_extended_4 = decon_nnls_extended[30:32,]
decon_nnls_extended_5 = decon_nnls_extended[33:35,]
decon_nnls_extended.m   = reshape2::melt(decon_nnls_extended)
decon_nnls_extended_0.m = reshape2::melt(decon_nnls_extended_0)
decon_nnls_extended_1.m = reshape2::melt(decon_nnls_extended_1)
decon_nnls_extended_2.m = reshape2::melt(decon_nnls_extended_2)
decon_nnls_extended_3.m = reshape2::melt(decon_nnls_extended_3)
decon_nnls_extended_4.m = reshape2::melt(decon_nnls_extended_4)
decon_nnls_extended_5.m = reshape2::melt(decon_nnls_extended_5)
decon_nnls_extended.m

```

Remove CAD samples and rename conditions
```{r Remove_CAD_samples_and_rename}
decon_nnls_extended_tgs.m = cbind(decon_nnls_extended.m,"treatment"= methylKit::getTreatment(meth.deT),"samples_id"= methylKit::getSampleID(meth.deT))

decon_nnls_extended_tgs.m <-decon_nnls_extended_tgs.m %>% filter(treatment != 4 & treatment != 5)

decon_nnls_extended_tgs.m$treatment_descr <-decon_nnls_extended_tgs.m$treatment

decon_nnls_extended_tgs.m$treatment_descr <-str_replace(decon_nnls_extended_tgs.m$treatment_descr,"0","Healthy")
decon_nnls_extended_tgs.m$treatment_descr <-str_replace(decon_nnls_extended_tgs.m$treatment_descr,"1","Stemi")
decon_nnls_extended_tgs.m$treatment_descr <-str_replace(decon_nnls_extended_tgs.m$treatment_descr,"2","Nstemi")
decon_nnls_extended_tgs.m$treatment_descr <-str_replace(decon_nnls_extended_tgs.m$treatment_descr,"3","UA")
decon_nnls_extended_tgs.m$treatment_descr <- factor(decon_nnls_extended_tgs.m$treatment_descr , levels=c("Stemi", "Nstemi", "UA","Healthy"))

saveRDS(decon_nnls_extended_tgs.m,"../Results/cardiac/RDS/decon_cell_proportions.RDS")
```

Merge metadata with cell type proportion results
```{r Merge_metadata_with_results}
tbl_2 <- suppl_tbl_2 %>%
  dplyr::select(Sample, cfDNA_ug_per_ml_plasma) %>%
  rename(
    samples_id = Sample
  )
decon_nnls_extended_tgs.m_normalized <- decon_nnls_extended_tgs.m
decon_nnls_extended_tgs.m_normalized <- left_join(decon_nnls_extended_tgs.m_normalized, tbl_2, by = "samples_id")
decon_nnls_extended_tgs.m_normalized
```

Plot results
```{r Plot_Results,fig.height=16,fig.width=10}
# bp_decon_nnls_extended <- ggplot2::ggplot(decon_nnls_extended_tgs.m, ggplot2::aes(x=reorder(variable,value), y=value, fill=factor(treatment_descr))) + ggplot2::geom_boxplot()
# 
# bp_decon_nnls_extended + ggplot2::theme(legend.text=element_text(size=14),axis.text.x = ggplot2::element_text(size=14),axis.text.y = ggplot2::element_text(size=14), plot.title = ggplot2::element_text(size=14)) + ggplot2::labs(fill="Treatment",x = "Cell Type", y = "Proportion of Sample") + coord_flip()
```

Exclude non-haematological cell types and normalised data by amount of ccfDNA
```{r}
decon_nnls_extended_tgs.m_normalized <-decon_nnls_extended_tgs.m_normalized %>% filter(variable != "Adipocytes" & 
                                         variable != "Lung_cells" & 
                                         variable != "Pancreatic_beta_cells" & 
                                         variable != "Pancreatic_acinar_cells"	 &
                                         variable != "Pancreatic_duct_cells" &
                                         variable != "Colon_epithelial_cells" &
                                         variable != "Head_and_neck_larynx" &
                                         variable != "Prostate" &
                                         variable != "Upper_GI" &
                                         variable != "Left_atrium" &
                                         variable != "right_atrium_auricular_region" &
                                         variable != "Breast" &
                                         variable != "Uterus_cervix"
                                         )

decon_nnls_extended_tgs.m_normalized$value = decon_nnls_extended_tgs.m_normalized$value * decon_nnls_extended_tgs.m_normalized$cfDNA_ug_per_ml_plasma
decon_nnls_extended_tgs.m_normalized$method <- "ccfDNA (μg/ml of plasma)"
```

Exclude non-haematological cell types
```{r Exclude_celltypes}
decon_nnls_extended_tgs.m_short <-decon_nnls_extended_tgs.m %>% filter(variable != "Adipocytes" & 
                                         variable != "Lung_cells" & 
                                         variable != "Pancreatic_beta_cells" & 
                                         variable != "Pancreatic_acinar_cells"	 &
                                         variable != "Pancreatic_duct_cells" &
                                         variable != "Colon_epithelial_cells" &
                                         variable != "Head_and_neck_larynx" &
                                         variable != "Prostate" &
                                         variable != "Upper_GI" &
                                         variable != "Left_atrium" &
                                         variable != "right_atrium_auricular_region" &
                                         variable != "Breast" &
                                         variable != "Uterus_cervix"
                                         )

decon_nnls_extended_tgs.m_short$value = decon_nnls_extended_tgs.m_short$value * 100
decon_nnls_extended_tgs.m_short$method <- "% of cell/tissue"

decon_nnls_extended_all_short<-rbind(decon_nnls_extended_tgs.m_short,decon_nnls_extended_tgs.m_normalized %>% dplyr::select(-cfDNA_ug_per_ml_plasma))
decon_nnls_extended_all_short$treatment_descr <- factor(decon_nnls_extended_all_short$treatment_descr , levels=c("Stemi", "Nstemi", "UA","Healthy"))

#specify cell types more specifically
decon_nnls_extended_all_short_even_shorter <-decon_nnls_extended_all_short %>% filter(variable =="Vascular_endothelial_cells" | variable =="heart_left_ventricle" | variable =="Erythrocyte_progenitors" | variable =="coronary_artery" | variable =="Monocytes_EPIC"  | variable == "CD4T.cells_EPIC" | variable == "NK.cells_EPIC" | variable =="Kidney" | variable == "Neutrophils_EPIC" | variable =="Hepatocytes")
decon_nnls_extended_all_short_even_shorter$variable <-str_replace(decon_nnls_extended_all_short_even_shorter$variable,"_EPIC","")
decon_nnls_extended_all_short_even_shorter$variable <-str_replace(decon_nnls_extended_all_short_even_shorter$variable,"_"," ")
decon_nnls_extended_all_short_even_shorter$variable <-str_replace(decon_nnls_extended_all_short_even_shorter$variable,".cells","")
decon_nnls_extended_all_short_even_shorter$variable <-str_replace(decon_nnls_extended_all_short_even_shorter$variable,"_"," ")
decon_nnls_extended_all_short_even_shorter

saveRDS(decon_nnls_extended_all_short,"decon_nnls_extended_all_short.RDS")
```

```{r}
decon_nnls_extended_all_short_even_shorter %>%
  mutate(
    variable = str_to_title(variable),
    treatment_descr = fct_recode(treatment_descr,
      "STEMI" = "Stemi",
      "NSTEMI" = "Nstemi"
    )
  )
```

# Plotting

```{r Publication Figure}
Figure_2 <-
  decon_nnls_extended_all_short_even_shorter %>%
  mutate(
    variable = case_when(
      variable == "coronary artery" ~ "Coronary artery",
      variable == "heart left ventricle" ~ "Heart left ventricle",
      variable == "CD4T" ~ "CD4+ T cells",
      variable == "NK" ~ "NK cells",
      variable == "Vascular endothelial" ~ "Vascular endothelial cells",
      TRUE ~ variable
    ),
    treatment_descr = fct_recode(treatment_descr,
      "STEMI" = "Stemi",
      "NSTEMI" = "Nstemi"
    )
  ) %>%
  filter(method == "% of cell/tissue") %>%
  ggplot(ggplot2::aes(x = variable, y = value, fill = factor(treatment_descr))) +
  geom_boxplot() +
  theme_Publication() +
  scale_fill_Publication() +
  theme(
    axis.text.x = element_blank(),
    strip.text = element_text(size = 12),
    legend.text = element_text(size = 14),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank()
  ) +
  labs(fill = "Group", x = "treatment_descr", y = "Predicted origin of ccfDNA [%]") +
  facet_wrap(~variable, scale = "free", ncol = 5, labeller = label_wrap_gen(width = 15), strip.position = "top")

ggsave(
  bg = "white",
  "/local/AAkalin_cardiac/Results/cardiac/Plots/Figure2.pdf",
  device = "pdf",
  plot = Figure_2, 
  width = 210, 
  units = "mm", 
  height = 180
)
Figure_2
```



```{r Plot_haematological_cell_types_shorter_without_y_label, fig.height=18,fig.width=12}
#pdf(file = "/local/AAkalin_cardiac/Results/cardiac/Plots/Figure2.pdf",height = 18,width = 12)


bp_decon_nnls_extended <- ggplot(decon_nnls_extended_all_short_even_shorter, aes(x=variable, y=value, fill=factor(treatment_descr))) + 
  geom_boxplot()

bp_decon_nnls_extended +   
  theme_Publication() + 
  scale_fill_Publication() + 
  theme(axis.text.x = element_blank(),
        strip.text = element_text(size=12),
        legend.text=element_text(size=14),
        axis.title.x = element_blank()) + 
  labs(fill="Group",x = "treatment_descr", y = "") + 
  facet_wrap(~variable + method, scale="free",ncol = 4) 

ggplot2::ggsave(filename = "/local/AAkalin_cardiac/Results/cardiac/Plots/Figure2.pdf", 
plot =  last_plot(), 
width = 12,
height = 18)

#dev.off()
```

```{r Plot_haematological_cell_types_reordered,fig.height=16,fig.width=12}
bp_decon_nnls_extended <- ggplot2::ggplot(decon_nnls_extended_all_short, ggplot2::aes(x=reorder(variable, value), y=value, fill=factor(treatment_descr))) + ggplot2::geom_boxplot()

bp_decon_nnls_extended + ggplot2::theme(legend.text=element_text(size=14),axis.text.x = ggplot2::element_text(size=14), axis.text.y = ggplot2::element_text(size=14),plot.title = ggplot2::element_text(size=16)) + ggplot2::labs(fill="Treatment",x = "", y = "") + coord_flip() + facet_wrap(~method,scales = "free_x")
```

```{r,fig.height=16,fig.width=10}
bp_decon_nnls_extended <- ggplot2::ggplot(decon_nnls_extended_tgs.m_short, ggplot2::aes(x=reorder(variable,value), y=value, fill=factor(treatment_descr))) + ggplot2::geom_boxplot()

bp_decon_nnls_extended + ggplot2::theme(legend.text=element_text(size=14),axis.text.x = ggplot2::element_text(size=14), axis.text.y = ggplot2::element_text(size=14),plot.title = ggplot2::element_text(size=20)) + ggplot2::labs(fill="Treatment",x = "Cell Type", y = "Proportion of Sample") + coord_flip()


ggsave(
  "../Results/cardiac/Plots/Figure2.pdf",
  device = "pdf",
  plot = last_plot())
```

```{r Previous_plot_without_theme}
bp_decon_nnls_extended <- ggplot2::ggplot(decon_nnls_extended_tgs.m_short, ggplot2::aes(x=variable, y=value, fill=factor(treatment_descr))) + ggplot2::geom_boxplot()
```

```{r Adding_theme,fig.height=20,fig.width=16}
bp_decon_nnls_extended + ggplot2::theme(axis.text.x = element_blank(),strip.text = element_text(size=16),legend.text=element_text(size=14),axis.title.x = element_blank()) + ggplot2::labs(fill="Treatment",x = "Cell Type", y = "Proportion of Sample") + ggplot2::facet_wrap(~variable, scale="free") + ggplot2::scale_y_continuous(expand = c(0, 0), limits = c(0, NA))
```

Investigate Neutrophils of healthy patients
```{r Check_healthy_patients}
neut_health<-decon_nnls_extended_tgs.m_short %>% filter(variable == "Neutrophils_EPIC" & treatment ==0) %>% dplyr::select(value)
mean(neut_health$value)

health_composition<-decon_nnls_extended_tgs.m %>% filter(treatment ==0) %>% group_by(variable) %>% summarise_at("value",mean)
```

