---
title: "Differentially Methylated  Regions (DMR) annotation and ranking"
output: html_notebook
---


## This report shows the results of DMR analysis comparing each disease group vs. healthy controls 

Run ranking_v3.rmd first.
The DMRs were called using methylKit R package, with adjusted q-value threshold 0.01 and minimal methylation difference 25%
The table shows the number of significant DMRs obtained for each group.


```{r echo=FALSE}
kbl(counting %>% dplyr::select("Number of DMRs"))
```


## Distribution of DMRs across diseases/groups
Comparing the DMRs across diseases, we found disease exclusive DMRs and shared DMRs between some groups. 
The figure shows a Upset Plot, displaying the number of DMRs exclusive and shared across all the groups.
We can see, for example, that from the 865 DMRs found for iAP vs. controls, 607 are exclusively found for this group. 
We also can see that Stemi and Nstemi share a total of 123 DMRs (60+63).
```{r echo=FALSE, fig.height = 6, fig.width = 10}
plt<-UpSetR::upset(UpSetR::fromList(v), order.by = "freq")
```

```{r echo=FALSE,fig.height = 6, fig.width = 10}
print(plt)
```



## DMR gene annotation
We used [GREAT](http://bejerano.stanford.edu/great/public/html/) to annotate the DMRs with genes. 
The following figures shows the number of genes per region and  the distance to transcription starting site (TSS) for each group.
We can see that for most of regions there are at least one gene annotated, but for most there are 2 and most of DMRs are 50 to 500 kb distant of the TSS

### STEMI
```{r echo=FALSE,fig.height = 6, fig.width = 12}
STEMI<-plotRegionGeneAssociationGraphs(geneList$ACS_Stemi$job, type=c(1,3))
```


### NSTEMI
```{r echo=FALSE,fig.height = 6, fig.width = 12}
NSTEMI<-plotRegionGeneAssociationGraphs(geneList$ACS_Nstemi$job, type=c(1,3))

```


### UA
```{r echo=FALSE,fig.height = 6, fig.width = 12}
UA<-plotRegionGeneAssociationGraphs(geneList$ACS_iAP$job, type=c(1,3))
```

```{r Place_Figure_Label}
fig_label <- function(text, region="figure", pos="topleft", cex=NULL, ...) {

  region <- match.arg(region, c("figure", "plot", "device"))
  pos <- match.arg(pos, c("topleft", "top", "topright", 
                          "left", "center", "right", 
                          "bottomleft", "bottom", "bottomright"))

  if(region %in% c("figure", "device")) {
    ds <- dev.size("in")
    # xy coordinates of device corners in user coordinates
    x <- grconvertX(c(0, ds[1]), from="in", to="user")
    y <- grconvertY(c(0, ds[2]), from="in", to="user")

    # fragment of the device we use to plot
    if(region == "figure") {
      # account for the fragment of the device that 
      # the figure is using
      fig <- par("fig")
      dx <- (x[2] - x[1])
      dy <- (y[2] - y[1])
      x <- x[1] + dx * fig[1:2]
      y <- y[1] + dy * fig[3:4]
    } 
  }

  # much simpler if in plotting region
  if(region == "plot") {
    u <- par("usr")
    x <- u[1:2]
    y <- u[3:4]
  }

  sw <- strwidth(text, cex=cex) * 60/100
  sh <- strheight(text, cex=cex) * 60/100

  x1 <- switch(pos,
    topleft     =x[1] + sw, 
    left        =x[1] + sw,
    bottomleft  =x[1] + sw,
    top         =(x[1] + x[2])/2,
    center      =(x[1] + x[2])/2,
    bottom      =(x[1] + x[2])/2,
    topright    =x[2] - sw,
    right       =x[2] - sw,
    bottomright =x[2] - sw)

  y1 <- switch(pos,
    topleft     =y[2] - sh,
    top         =y[2] - sh,
    topright    =y[2] - sh,
    left        =(y[1] + y[2])/2,
    center      =(y[1] + y[2])/2,
    right       =(y[1] + y[2])/2,
    bottomleft  =y[1] + sh,
    bottom      =y[1] + sh,
    bottomright =y[1] + sh)

  old.par <- par(xpd=NA)
  on.exit(par(old.par))

  text(x1, y1, text, cex=cex, ...)
  return(invisible(c(x,y)))
}
```

```{r Export_Supplement_S4}
condition= c("STEMI","NSTEMI" ,"UA", "CAD", "CADpos", "CADneg")
labels = c("A", "B", "C")

for(i in 1:3){
    job=geneList[[i]]$job
  pdf(paste0("/local/AAkalin_cardiac/Results/cardiac/Plots/Figure_S4", labels[i], ".pdf"), width = 10, height = 5)
  par(xpd=NA)
  plotRegionGeneAssociationGraphs(job, type=c(1,3))
  title(main = paste0(condition[i]), cex.main = 2,   font.main= 2, line = 2.8)  
  fig_label(labels[i], region="figure", pos="topleft", cex=3, font=2)
  dev.off()
}
```
### CAD+posStressEcho+pos
```{r echo=FALSE,fig.height = 6, fig.width = 12}
CADpos<-plotRegionGeneAssociationGraphs(geneList$`CAD+posStressEcho+pos`$job, type=c(1,3))
```

### CAD+posStressEcho+neg
```{r echo=FALSE,fig.height = 6, fig.width = 12}
CADneg<-plotRegionGeneAssociationGraphs(geneList$`CAD+posStressEcho+neg`$job, type=c(1,3))
```

## DMRs Gene-disease associations
We used [DisGeNET](https://www.disgenet.org/) to find gene-disease associations for the DMRs and then we filted for those related to heart disases.
The following table shows the number of DMRs with heart-related-disease associations found and also the same for DMRs group specific (disease exclusive). We can
see that for MI groups the number of DMRs annotated with genes associated with heart related diseases are the largest.
```{r echo=FALSE}
kbl(counting %>% dplyr::select("Number of DMRs with gene-disease associations", "Number of DMRs disease exclusive with gene-disease associations"))
```


The figure shows a Upset Plot, displaying the number of DMRs exclusive and shared across all the groups, but now just for the DMR with genes associated to heart-related diseases.
```{r echo=FALSE, fig.height = 6, fig.width = 10}
plt_2<-UpSetR::upset(UpSetR::fromList(v_e), order.by = "freq")
```

```{r echo=FALSE,fig.height = 6, fig.width = 10}
print(plt_2)
```
We investigated also Disease enrichments also from DisGeNET. The enrichment is filtered by adjusted p-values < 0.05.


### ACS_Stemi
```{r echo=FALSE, fig.height = 6, fig.width = 12}
print(enrichment_all$ACS_Stemi)
```

### ACS_Nstemi
```{r echo=FALSE, fig.height = 6, fig.width = 12}
print(enrichment_all$ACS_Nstemi)
```

### ACS_iAP
```{r echo=FALSE, fig.height = 6, fig.width = 12}
print(enrichment_all$ACS_iAP)
```

### CAD+posStressEcho+pos
```{r echo=FALSE, fig.height = 6, fig.width = 12}
print(enrichment_all$`CAD+posStressEcho+pos`)
```


### CAD+posStressEcho+neg
```{r echo=FALSE, fig.height = 6, fig.width = 12}
print(enrichment_all$`CAD+posStressEcho+neg`)
```




The tables with all DMRs statistics and annotations are shown below. 
The DRMs on the tables are ranked (sorted) by absolute meth.diff (effect size) ---> qvalue -> SD (within group variation) ---> heart_disease_related (Y|N) --> max_score (max score of heart related diseases for that DMR) --> max_ei ---> max_dsi
## STEMI
```{r echo=FALSE}
kbl(final_DMR_df_SDs_sorted$ACS_Stemi %>% dplyr::select("seqnames","start","end","width","strand","gene" ,"patent_genes","circulation_genes_MI",
                                                        "circulation_genes_CHD","exclusive","present_in_groups",
                                                        "heart_disease_related_disgenet","heart_disease_name_disgenet",
                                                        "heart_disease_class_name_disgenet","EFO_id_heart_related_disgenet",
                                                        "max_score_disgenet","max_ei","max_dsi","EFO0001645_score_open_target",
                                                        "meth.diff","qvalue","SD","SD_control")) %>%
                                                         kable_paper() %>%
                                                         scroll_box(width = "100%", height = "400px")
```


## NSTEMI
```{r echo=FALSE}

kbl(final_DMR_df_SDs_sorted$ACS_Nstemi %>% dplyr::select("seqnames","start","end","width","strand","gene" ,"patent_genes","circulation_genes_MI",
                                                        "circulation_genes_CHD","exclusive","present_in_groups",
                                                        "heart_disease_related_disgenet","heart_disease_name_disgenet",
                                                        "heart_disease_class_name_disgenet","EFO_id_heart_related_disgenet",
                                                        "max_score_disgenet","max_ei","max_dsi","EFO0001645_score_open_target",
                                                        "meth.diff","qvalue","SD","SD_control")) %>%
                                                         kable_paper() %>%
                                                         scroll_box(width = "100%", height = "400px")
```


## ACS_iAP
```{r echo=FALSE}
kbl(final_DMR_df_SDs_sorted$ACS_iAP %>% dplyr::select("seqnames","start","end","width","strand","gene" ,"patent_genes","circulation_genes_MI",
                                                        "circulation_genes_CHD","exclusive","present_in_groups",
                                                        "heart_disease_related_disgenet","heart_disease_name_disgenet",
                                                        "heart_disease_class_name_disgenet","EFO_id_heart_related_disgenet",
                                                        "max_score_disgenet","max_ei","max_dsi","EFO0001645_score_open_target",
                                                        "meth.diff","qvalue","SD","SD_control")) %>%
                                                         kable_paper() %>%
                                                         scroll_box(width = "100%", height = "400px")
```



## CAD+posStressEcho + Pos
```{r echo=FALSE}
kbl(final_DMR_df_SDs_sorted$'CAD+posStressEcho+pos' %>% dplyr::select("seqnames","start","end","width","strand","gene" ,"patent_genes","circulation_genes_MI",
                                                        "circulation_genes_CHD","exclusive","present_in_groups",
                                                        "heart_disease_related_disgenet","heart_disease_name_disgenet",
                                                        "heart_disease_class_name_disgenet","EFO_id_heart_related_disgenet",
                                                        "max_score_disgenet","max_ei","max_dsi","EFO0001645_score_open_target",
                                                        "meth.diff","qvalue","SD","SD_control")) %>%
                                                         kable_paper() %>%
                                                         scroll_box(width = "100%", height = "400px")

```

## CAD+posStressEcho + Neg
```{r echo=FALSE}
kbl(final_DMR_df_SDs_sorted$'CAD+posStressEcho+neg' %>% dplyr::select("seqnames","start","end","width","strand","gene" ,"patent_genes","circulation_genes_MI",
                                                        "circulation_genes_CHD","exclusive","present_in_groups",
                                                        "heart_disease_related_disgenet","heart_disease_name_disgenet",
                                                        "heart_disease_class_name_disgenet","EFO_id_heart_related_disgenet",
                                                        "max_score_disgenet","max_ei","max_dsi","EFO0001645_score_open_target",
                                                        "meth.diff","qvalue","SD","SD_control")) %>%
                                                         kable_paper() %>%
                                                         scroll_box(width = "100%", height = "400px")
```


```{r echo=FALSE}
saveRDS(final_DMR_df_SDs_sorted,"final_DMR_df_SDs_sorted.RDS")
```

